Modeling memory and taxis to analyze migrations

### 2021-06-23 text/ms_v2.Rmd
ms ms ms ms ms

The drifting resource differs from the island resource in that the total amount of resource at any given time point is 1 ($\int_X h(x,t) dx = 1$).  This property is attained by distributing the resource as a re-scaled beta distribution, where the shape and scale parameters vary sinusoidally in such a way as to make the standard deviations and means match the desired values of $m_x, m_t, s_x, s_t$ (see appendix for details).  Both types of resources are illustrated in Figure 1.

@@

@@

An illustration of the ability to adjust its migration to find a stable, migratory state that matches the dynamics of the resource is illustrated in Figure X.  Under many combinations of parameters, this kind of steady state adjustment is obtained and maintained.

@@

In general, greater values of $\alpha$ (resource following) lead to an improved ability to match the migration,

Resource peaks with larger spatial extent (high $\sigma_x$) are smaller duration (low $\sigma_t$) are generally better for migration matching, suggesting that broad but short pulses have the strongest impact on the migration memory.

Larger values of social attraction ($\beta$) generally lead to poorer migration matching except in those cases where the sociality scale $\lambda$ was high.  Thus, for example, at $\lambda = 20$, no simulations at $\beta >= 200$ managed to acquire or maintain a matched migration.  However, at $\lambda = 50$ or $100$, the migration was slightly better matched at high values of $\beta.  (Fig. **X**)

Higher diffusion rates

@@

![](figures/example1_adaptation.png)
**Figure 1. Example of adaptation to a shift in resource peak. The initial (year 0) behavior migrates to locations 50 and -50 at days 15 and 60, whereas the resource peak is at 30 and -30, peaking at times 25 and 75.  (a) shows the first 14 years of the simulation; (b) The centroid of the annual movement of the population is shown in panel b, with dark blue to red colors indicating year 0 to year 40; (c) annual foraging efficiency across years; (d) migration timing parameters for each year, with orange segments indicating arrival and departure from the summering (northern) grounds, and the blue segments indicating timing of arrival and departure at the wintering grounds; (e) migration arrival and departure location across years, with blue and orange indicating winter (southern) and summer (northern) locations.**

>>>>

The drifting resource differs from the island resource in that the total amount of resource at any given time point is 1 ($\int_X h(x,t) dx = 1$).  This property is attained by distributing the resource as a re-scaled beta distribution, where the shape and scale parameters vary sinusoidally in such a way as to make the standard deviations and means match the desired values of $m_x, m_t, s_x, s_t$ (see appendix for details).

@@

![](figures/ResourceExamples.png)
**Figure 1. Examples of various seasonal resource distribution functions, contrasting short duration, but wide pulses ($\sigma_t = 3, \sigma_x = 12$; left panels), long duration but spatially concentrated pulses ($\sigma_t = 12, \sigma_x = 3$; right panels), and isolated resource pulses (upper panels) from the weakly drifting resource (lower panels).  The total amount of resource is identical across all scenarios. In the weakly drifting resources, the total amount is constant at all times, and uniform in the middle of the phase (time = 0, 50, 100).**

@@

sigma_t, sigma_x, 
epsilon, alpha, 
beta, lambda, 
x_peak = 30, t_peak = 25

@@

- In general, greater values of $\alpha$ (resource following) lead to an improved ability to match the migration,

- Resource peaks with larger spatial extent (high $\sigma_x$) are smaller duration (low $\sigma_t$) are generally better for migration matching, suggesting that broad but short pulses have the strongest impact on the migration memory.

- Larger values of social attraction ($\beta$) generally lead to poorer migration matching except in those cases where the sociality scale $\lambda$ was high.  Thus, for example, at $\lambda = 20$, no simulations at $\beta >= 200$ managed to acquire or maintain a matched migration.  However, at $\lambda = 50$ or $100$, the migration was slightly better matched at high values of $\beta.  (Fig. **X**)

- Higher diffusion rates

@@

![](figures/example1_adaptation.png)
**Figure 2. Example of adaptation to a shift in resource peak. The initial (year 0) behavior migrates to locations 50 and -50 at days 15 and 60, whereas the resource peak is at 30 and -30, peaking at times 25 and 75.  (a) shows the first 14 years of the simulation; (b) The centroid of the annual movement of the population is shown in panel b, with dark blue to red colors indicating year 0 to year 40; (c) annual foraging efficiency across years; (d) migration timing parameters for each year, with orange segments indicating arrival and departure from the summering (northern) grounds, and the blue segments indicating timing of arrival and departure at the wintering grounds; (e) migration arrival and departure location across years, with blue and orange indicating winter (southern) and summer (northern) locations.**
